### 2022-12-07 paper.Rmd
Abstract

    Anticlustering is a collection of methods to partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximzing instead of minimizing a clustering objective function. Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications. While k-means anticlustering minimizes between-group differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize differences with regard to several additional distribution characteristics, including the variance, skewness and kurtosis, and even covariance structure. It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria. In particular, optimizing similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering. Higher order moments and covariance structure can also succesfully be equalized, while fulfilling multiple criteria at the same time is improved when the sample size increases. Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, higher order moments"
wordcount         : "X"

>>>>

    Anticlustering is a collection of methods to partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---and is usually approached by maximzing instead of minimizing a clustering objective function. Introducing k-plus, this paper presents an extension of the k-means clustering objective to maximize between-group similarity in anticlustering applications. The k-plus criterion represents between-group similarity as discrepancy in distribution moments (means, variance, higher order moments, and covariances), thereby extending the classical k-means criterion, which only reflects how groups differ in their means. It is shown that the k-plus criterion can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple objectives. In particular, optimizing between-group similarity with regard to variances as well as means comes with almost no cost to either criterion; the k-plus extension is therefore generally preferred over classical k-means anticlustering. Higher order moments and covariance structure can also succesfully be equalized. Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis"
wordcount         : "X"
